---
title: "MAT iris"
author: "Sofian Hamiti"
output: pdf_document
---

#Introduction
The Iris dataset contains 150 instances, corresponding to three 
equally-frequent species of iris plant (Iris setosa, Iris versicolour, and Iris virginica).

This project has been built in 2 parts, using R:

- This presentation document 

- A ShinyApp to retrieve the 10 most similar observations to the inputted vector


#The Data
## Loading the data:
```{r, message = FALSE, results='hide', warning=FALSE}
library("ggplot2")
library("gridExtra")
source("cleandata.R")

#Load the dataset from the website and assign names to columns
iris <- read.csv(url("http://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data")) 
colnames(iris) <- c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species")
```

##Simple data quality assessment
```{r}
#Number of NA values in the dataset
sum(is.na(iris))

#Show the duplicate values
subset(iris, duplicated(iris))

#Call the cleanData function in order to remove NAs, Duplicates, and Negative values.
iris <- cleanData(iris)
```

##Analysis of the Tidy dataset
```{r}
#Number of observations and the attributes' data types
str(iris)
#Distribution of species
table(iris$Species)
#Summary statistics of the flowers' specs from which 
#I have defined the extreme values to input in the shinyapp.
summary(iris[,-5])
```


#Exploratory analysis
```{r}
#As we can see in the charts, in future developments, it may be interesting to 
#cluster the observations using their Petal specs.
```
```{r, echo=FALSE}
plot_sepal <- ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, colour=Species)) +
        geom_point() +
        ggtitle("Distribution of Sepal.Length~Sepal.Width")
plot_petal <- ggplot(iris, aes(x=Petal.Length, y=Petal.Width, colour=Species)) +
        geom_point() +
        ggtitle("Distribution of Petal.Length~Petal.Width")
grid.arrange(plot_sepal, plot_petal, nrow=2)
```


#The ShinyApp
The app has been built using R and has the following functionalities:

- Reads in the data, accessed from: https://archive.ics.uci.edu/ml/datasets/Iris;

- Clean the data  using the cleanData function.

- Takes as arguments, attributes defining a new species of Iris plant with the use of sliders. 
        Sliders helped me to protect the app from unrealistic inputs.

- Returns the ten most similar data points in the existing Iris data based on the euclidean distance from the inputted vector.

- Displays the results in a table.

The code is available on my github repository:
https://github.com/SofianHamiti/MAT_iris

The ShinyApp:
https://sofianhamiti.shinyapps.io/MAT_iris

#Future Possible Improvements

- Improve the accuracy of the calculation. Cosine Similarity may be a good start of research.

- Visualize the output in a plot to have a more intuitive vision of the distance calculated.
